CrypTool Script\footnote{ Background reading for the CrypTool program}\\ Mathematics and Cryptography

### 2003-04-07 script/en/chapters/numbertheory.tex
richard's comments

Although these considerations and definitions may seem to be rather theoretical, we are so familiar with them in 
everyday life that we no longer think about the formal procedure. For example, the 24 hours on a clock are 
represented by the numbers $1, 2, \cdots, 12$. We obtain the hours after 12 noon as the remainder of a division by 12 
and know immediately that 2 o’clock in the afternoon is the same as 14.00.

@@

A calculation result larger than 4 is then transformed ``modulo $5$''. In other words, it is the remainder when the 
result is divided by $5$. For example, $2*4 \equiv 8 \equiv 3 {\rm ~(mod~ } 5)$ because $3$ is the remainder 
when we divide $8$ by $5$.

@@

Cryptographically not only the uniqueness of the inverse is important, but also that the
set of possible values is exhausted.
\begin{theorem}\label{thm-zth-exhperm}
For $a,i\in \{1, \cdots, m-1 \}$ with ${\rm~gcd} (a,m)=1),$ then takes the product $a*i {\rm ~mod ~} m $ for a 
certain number $a$ all values from $\{1, \cdots ,m-1 \} $ (exhaustive permutation \index{Permutation} of the length $m-1$)\footnote{See
also theorem \ref{thm-zth-ordp} in \hyperlink{Kap_3_9}{chapter 3.9 Multiplicative order and primitive roots}.}.
\end{theorem}
\vskip +10 pt
The following three examples\footnote{%
See \hyperlink{AppArith1}{Appendix D} for the source code to compute
the tables using Mathematica\index{Mathematica} and Pari-GP.\index{Pari-GP}
}
illustrate the properties of multiplicative inverses.

@@

{\bf Example 1:}
$$ (4^3)^2 \equiv 64^2 \equiv 4096 \equiv 1 {\rm ~(mod~} 5). $$
\begin{quote}
(1) We can speed up\footnote{The time required to calculate the multiplication 
of two numbers normally depends on the length of the numbers.  We can observe
this if we use the school method to calculate, for instance, $474*228$.  The 
time required increases 
%quadratically
in a quadratic square manner , because we need to multiply $3*3$ numbers. The 
numbers become considerably smaller if we reduce the interim result.} the 
calculation by reducing the {\bf interim results} modulo $5$ but we need 
to take care because not everything then works in the same way as in 
standard arithmetic.
\begin{eqnarray*}
(4^3)^2 & \equiv & (4^3{\rm ~(mod~}5))^2{\rm ~(mod~}5) \nonumber \\
            & \equiv & (64{\rm ~(mod~}5))^2{\rm ~(mod~}5) \nonumber \\
            & \equiv & 4^2{\rm ~(mod~}5) \nonumber \\
            & \equiv & 16 \equiv 1 {\rm ~(mod~}5). \nonumber
\end{eqnarray*}

@@

If we know $a$ and $b$ of the three numbers $a$, $b$ and $c$ that meet this equation, then every known method 
of finding $c$ is approximately just as time-consuming as trying out all $m$ possible values for $c$ one after the 
other. For a typical $m$ of the order of magnitude of $10^{180}$ for $600$-digit binary numbers, this is a 
hopeless undertaking. More precisely, for suitably large numbers $m$, the time required according to current 
knowledge is proportional to ${\rm exp}\left( C*( \log m [\log \log m]^2)^{1/3}\right)$ with a constant $C > 1$.
\item[\bf b)] {\bf Calculating the root (determining $b$):}

@@

If $J(m)$ cannot be determined\footnote{According to the first fundamental theorem of number 
theory\index{Number theory!fundamental theorem} and theorem \ref{thm-zth-phinum}, we determine $J(m)$ by reducing $m$ to prime factors\index{Prime factor!decomposition}.}, it is difficult to calculate 
the $c$-th root. This forms the basis for the security assumption used by the RSA encryption system (see 
Sub-section 4.3.1: \hyperlink{RSAVerfahren}{the RSA procedure} and \hyperlink{RSABeweis}{Sub-section 3.10 Proof of the RSA procedure}).
\end{itemize}
The time required for inverting addition and multiplication, on the other hand, 
is simply proportional to $\log m$ or $(\log m)^2$. Powers (for a number $x$ 
calculate $x^a$ with $a$ fixed) and exponents (for a number $x$ 
calculate $a^x$ with $a$ fixed) are therefore typical one way functions 
(See Overview of the one way functions in this \hyperlink{OneWayFunktion1}{Script} and
\hyperlink{OneWayFunktion2}{artikel}). \index{One way function}

@@

If we define the operation mod+ on such a set where
$$ a {\rm ~mod+~} b := (a + b){\rm ~(mod~}n) , $$
then the set $\mathbb{Z}_n$ together with the relation mod+ is a group because the following properties 
of a group are valid for all elements in $\mathbb{Z}_n$:
\begin{itemize}
\item   $ a {\rm ~mod+~} b$ is an element of $\mathbb{Z}_n$  (the set is closed),
\item   $(a {\rm ~mod+~} b) {\rm ~mod+~} c \equiv a {\rm ~mod+~} (b {\rm ~mod+~} c)$~~~  (mod+ is associative),
\item   the neutral element is $0$.
\item   each element $a \in \mathbb{Z}_n$ has an inverse for this operation, namely $n-a$  \\
        (because $a {\rm ~mod+~} (n-a) \equiv a + (n-a){\rm ~(mod~}n) \equiv n \equiv 0 {\rm ~(mod~}n)$).
\end{itemize}
Since the operation is commutative, i.e. $(a {\rm ~mod+~} b) = (b {\rm ~mod+~} a)$, this structure \index{Structure} is even a ``commutative group''.

@@

Mathematicians often ask, under which conditions the repeated application of an operation results in the neutral
element (compare patterns and structures before)\index{Structure}.

@@

\hypertarget{RSABeweis} Using the Euler-Fermat theorem, we can ``prove'' the
RSA\footnote{The RSA procedure\index{RSA procedure} is the most common asymmetric 
cryptography procedure. Developed in 1978 by Ronald Rivest, Adi Shamir and Leonard Adleman, it can be used both for
signatures and for encryption. Cryptographers always associate this procedure with the abbreviation ``RSA''. However, 
please note that the abbreviation may be used in other fields to mean
something completely different.}$^,$\footnote{%
Within literature and movies not only classical methods are used (like the
secret message solved by Sherlock Holmes in ``The Adventure of the Dancing
Men'' by Arthur Conan  Doyle \index{Arthur Conan Doyle}), but also modern schemes. 
Examples are
% hier die items nicht einrücken ! xxxxxxxxxxxxxxx
\begin{list}{\textbullet}{\leftmargin10pt}
\item the movie ``House of Cards'', 1992, where autistic children
  communicate via primes,
\item the story ``The Dialogue of the Sisters'' by Dr.~C.~Elsner\index{Elsner 1999}, 1999
  (included in the CrypTool package\index{CrypTool} as PDF file), where the sisters
  communicate confidentially using a variant of RSA.
\end{list}\vspace{-\baselineskip}\vspace{-\parskip}}
procedure in the group $\mathbb{Z}_n^*$.

@@

Although Whitfield Diffie and Martin Hellman formulated the general method as early as 1976, an actual procedure that 
met all four requirements was only found later by Rivest, Shamir and Adleman.

@@

Since this is not true in general, we need to consider the case when $M$ and $p$ are not relatively prime. Since $p$ is a 
prime number, this implies that $p$ is a factor of $M$. But this means:
$$  M \equiv 0 {\rm ~(mod~}p). $$
If $p$ is a factor of $M$, then $p$ is also a factor of $M^{e * d}$. Therefore:
$$M^{e * d} \equiv 0 {\rm ~(mod~}p).$$
Since $p$ is a factor of both $M$ and $Me * d$, it is also a factor of their difference:
$$ (M^{e * d} - M ) \equiv 0 {\rm ~(mod~}p).$$
And thus our conjecture is also true in this special case.

@@

{\bf 1st comment:} \\
We can also condense the three steps if we use theorem \ref{thm-zth-fermateuler} (Euler-Fermat) - i.e. not the simplified theorem where $n = p$ and 
which corresponds to Fermat's Little Theorem:
$$
(M^e)^d \equiv M^{e*d} \equiv M^{(p-1)(q-1)*k + 1} \equiv 
        (\underbrace{M^{(p-1)(q-1)}}_{\equiv M^{J(n)} \equiv 1 {\rm ~(mod~}n)})^k * M
    \equiv 1^k * M \equiv M {\rm ~(mod~}n).
$$

@@

This made evident, that a module length of 512 bit no longer prevents from
attackers.

@@

% --------------------------------------------------------------------------
\vskip +40pt
\subsubsection{Further actual research in factorisation and number theory}
Prime numbers are part of many very topical research areas in number theory and computer science. 
Progress made with factorisation is bigger than estimated 5 years ago -- this
is not only founded on faster computers but also in new knowledge.

@@

Thus, any progress in efficiency of factorizing large integers will effect
the security of the RSA. As a consequence, the underlying primes $p$, $q$
and, thus, the module n (1024 bit as off today) have to be increased. In
the worst case of a breakthrough in factorization, the RSA algorithm might
be compromised.

% --------------------------------------------------------------------------
\vskip +20pt
\paragraph{Bernstein's paper and it's implication on the
  security of the RSA algorithm}
\label{RSABernstein} \index{Factorisation problem}\mbox{}

@@

We note that the definition of \emph{effort} is a crucial point: Bernstein
claims that effort is the product of time and costs of the machine
(including the memory used). The gist of the paper lies in the fact that he
can reduce a big part of factorizing to sorting. Using Schimmler's scheme,
sorting can be optimized by massive parallel computing.  At the end of
section 3 Bernstein explains this effect: The costs of $m^2$ parallel
computers with a constant amount of memory is a constant times $m^2$.  The
costs of a computer with a single processor and memory of size $m^2$ is
also of the order of $m^2$, but with a diffrent constant factor.  With
$m^2$ processors in parallel, sorting of $m^2$ numbers (with Schimmler's
scheme) can be achieved in time $m$, while a $m^2$-memory computer needs
time of the order of $m^2$. Decreasing memory and increasing the number of
processors, one can reduce the computing time by a factor $1/m$ without
additional effort in terms of total costs.  In section 5 it is said that
massive parallel computing can also increase efficiency of factorizing
using Lenstra's elliptic-curve-method (a search algorithm has costs that
increase quadratically instead of cubically).

We note that all results achieved so far are asymptotic results. This means
that they only hold in the limit n to infinity. Unfortunately, there are no
upper bounds for the residual error (i.e. the difference between the real
and the asymptotic value) for finite n --- a problem which has already been
addressed by the author. As a consequence, one cannot conclude whether the
costs (in the sense of Bernstein) for factorizing 1024$-$2048-bit RSA modules
can be significantly reduced.

@@

When this section was written (June 2002) nothing was publicly known,
how far there exist implementations of his theoretical onsets and how
far financing of his research project was.

@@

Their abstract summarizes their results very well: ``The security of the RSA cryptosystem depends on the difficulty of factoring large integers. The best current factoring algorithm is the Number Field Sieve (NFS), and its most difficult part is the sieving step. In 1999 a large distributed computation involving thousands of workstations working for many months managed to factor a 512-bit RSA key, but 1024-bit keys were believed to be safe for the next 15-20 years. In this paper we describe a new hardware implementation of the NFS sieving step ... which is 3-4 orders of magnitude more cost effective than the best previously published designs ... . Based on a detailed analysis of all the critical components (but without an actual implementation), we believe that the NFS sieving step for 1024-bit RSA keys can be completed in less than a year by a \$10M device, and that the NFS sieving step for 512-bit RSA keys can be completed in less than ten minutes by a \$10K device. Coupled with recent results about the difficulty of the NFS matrix step ... this raises some concerns about the security of these key sizes.''

@@

There are three tasks with completely different degrees of difficulty here. In each case we know the ciphe rtext 
and the public key $(e,n)$:
\begin{itemize}
\item[{\bf (a)}] Known plaintext: \index{Attack!known plaintext} find the secret key $d$ using the additionally known original message.
\item[{\bf (b)}] Cipher text only: \index{Attack!cipher text only} find $d$ and the plaintext.
\item[{\bf (c)}] Calculate the RSA modulus, in other words factorisation (with no knowledge of the message). \index{Factorisation problem}
\end{itemize}

@@

\bibitem[Eckert2003]{Eckert2003} \index{Eckert 2003}
    Claudia Eckert, \\
    {\em IT-Sicherheit: Konzepte-Verfahren-Protokolle}, 
    Oldenbourg 2001, 2nd edition 2003.

@@

   \item \index{Munchenbach@M\"unchenbach}
         Only after I had completed this article, I did come across the
         website of Mr.\ M\"unchenbach, which interactively and didactically
         uses elementary number theory to provide a sophisticated
         description of the fundamental mathematical thought processes. 
         It was created for a teaching project in the 11th grade 
         of the technical grammar school (unfortunately only available in German): \\
         \href{http://www.hydrargyrum.de/kryptographie}
         {\texttt{http://www.hydrargyrum.de/kryptographie}}

    \item Also only after I had completed this article, I did come across
          the web site of Mr.~Wagner, who is responsible for the development of
          the curriculum of computer science in a German federal state. Here
          you can get (but only in German) a collection of texts and
          (Java-)\discretionary{}{}{}programs:\\
          \href{http://www.hom.saar.de/~awa/kryptolo.htm}
          {\texttt{http://www.hom.saar.de/\~{}awa/kryptolo.htm}}

@@

I would like to take this opportunity to thank 
\begin{itemize}
  \item {Mr. Henrik Koy for making many very useful suggestions, 
         for very constructive proof-reading this article and
         for helping with TeX. \\
	 Mr. Koy developed the functions and the complex dialog box of the
	 RSA cryptosystem, which allows you to execute the RSA samples
	 of this article.}
  \item {J\"org-Cornelius Schneider for his TeX support und for the many 
         cases where he helped when facing programming or design problems.}
  \item {Dr. Illies for pointing me to Pari-GP \index{Pari-GP}.}
\end{itemize}

@@

Loops are also supported by Pari-GP. Dispensing with the table formatting
the result looks like this:
\begin{verbatim}
for( x=1,12, 
     for(y=1,12, print(Mod(x^y,m))); 
     if(gcd(x,m)==1, print(znorder(Mod(x,m))), print("--"));
     print(eulerphi(m)))
\end{verbatim}

>>>>

Although these considerations and definitions may seem to be rather theoretical, we are so familiar with them in 
everyday life that we no longer think about the formal procedure. For example, the 24 hours on a clock are 
represented by the numbers $1, 2, \cdots, 12$. We obtain the hours after 12 noon as the remainder of a division by 12 
and know immediately that 2 o'clock in the afternoon is the same as 14.00.

@@

A calculation result larger than 4 is then reduced ``modulo $5$''. In other words, it is the remainder when the 
result is divided by $5$. For example, $2*4 \equiv 8 \equiv 3 {\rm ~(mod~ } 5)$ because $3$ is the remainder 
when we divide $8$ by $5$.

@@

Cryptographically not only the unique nature of the inverse is important, but
also that the set of possible values has been exhausted.
\begin{theorem}\label{thm-zth-exhperm}
For $a,i\in \{1, \cdots, m-1 \}$ with ${\rm~gcd} (a,m)=1),$ then takes the product $a*i {\rm ~mod ~} m $ for a 
certain number $a$ all values from $\{1, \cdots ,m-1 \} $ (exhaustive permutation \index{Permutation} of the length $m-1$)\footnote{See
also theorem \ref{thm-zth-ordp} in \hyperlink{Kap_3_9}{chapter 3.9 Multiplicative order and primitive roots}.}.
\end{theorem}
\vskip +10 pt
The following three examples\footnote{%
See \hyperlink{AppArith1}{Appendix D} for the source code to compute
the tables using Mathematica\index{Mathematica} and Pari-GP.\index{Pari-GP}
}
illustrate the properties of multiplicative inverses.

@@

{\bf Example 1:}
$$ (4^3)^2 \equiv 64^2 \equiv 4096 \equiv 1 {\rm ~(mod~} 5). $$
\begin{quote}
(1) We can speed up\footnote{The time required to calculate the multiplication 
of two numbers normally depends on the length of the numbers.  We can observe
this if we use the school method to calculate, for instance, $474*228$.  The 
time required increases 
%quadratically
in a quadratic square manner , because we need to multiply $3*3$ numbers. The 
numbers become considerably smaller if we reduce the interim result.} the 
calculation by reducing the {\bf interim results} modulo $5$ but we need 
to take care because not everything will then work in the same way as in 
standard arithmetic.
\begin{eqnarray*}
(4^3)^2 & \equiv & (4^3{\rm ~(mod~}5))^2{\rm ~(mod~}5) \nonumber \\
            & \equiv & (64{\rm ~(mod~}5))^2{\rm ~(mod~}5) \nonumber \\
            & \equiv & 4^2{\rm ~(mod~}5) \nonumber \\
            & \equiv & 16 \equiv 1 {\rm ~(mod~}5). \nonumber
\end{eqnarray*}

@@

If we know $a$ and $b$ of the three numbers $a$, $b$ and $c$ that meet this equation, then every known method 
of finding $c$ is approximately just as time-consuming as trying out all $m$ possible values for $c$ one after the 
other. For a typical $m$ of the order of magnitude of $10^{180}$ for $600$-digit binary numbers, this is a 
hopeless task. More precisely, for suitably large numbers $m$, the time required according to current 
knowledge is proportional to ${\rm exp}\left( C*( \log m [\log \log m]^2)^{1/3}\right)$ with a constant $C > 1$.
\item[\bf b)] {\bf Calculating the root (determining $b$):}

@@

If $J(m)$ cannot be determined\footnote{According to the first fundamental theorem of number 
theory\index{Number theory!fundamental theorem} and theorem \ref{thm-zth-phinum}, we determine $J(m)$ by reducing $m$ to prime factors\index{Prime factor!decomposition}.}, it is difficult to calculate 
the $c$-th root. This forms the basis for the security assumption used by the RSA encryption system (see 
Sub-section 4.3.1: \hyperlink{RSAVerfahren}{the RSA procedure} and \hyperlink{RSABeweis}{Sub-section 3.10 Proof of the RSA procedure}).
\end{itemize}
The time required for inverting addition and multiplication, on the other hand, 
is simply proportional to $\log m$ or $(\log m)^2$. Powers (for a number $x$ 
calculate $x^a$ with $a$ fixed) and exponents (for a number $x$ 
calculate $a^x$ with $a$ fixed) are therefore typical one way functions 
(See Overview of the one way functions in this \hyperlink{OneWayFunktion1}{Script} and
\hyperlink{OneWayFunktion2}{article}). \index{One way function}

@@

If we define the operation mod+ on such a set where
$$ a {\rm ~mod+~} b := (a + b){\rm ~(mod~}n) , $$
then the set $\mathbb{Z}_n$ together with the relation mod+ is a group because the following properties 
of a group are valid for all elements in $\mathbb{Z}_n$:
\begin{itemize}
\item   $ a {\rm ~mod+~} b$ is an element of $\mathbb{Z}_n$  (the set is closed),
\item   $(a {\rm ~mod+~} b) {\rm ~mod+~} c \equiv a {\rm ~mod+~} (b {\rm ~mod+~} c)$~~~  (mod+ is associative),
\item   the neutral element is $0$.
\item   each element $a \in \mathbb{Z}_n$ has an inverse for this operation, namely $n-a$  \\
        (because $a {\rm ~mod+~} (n-a) \equiv a + (n-a){\rm ~(mod~}n) \equiv n \equiv 0 {\rm ~(mod~}n)$).
\end{itemize}
Since the operation is commutative, i.e. $(a {\rm ~mod+~} b) = (b {\rm ~mod+~}
a)$, this structure \index{Structure} is actually a ``commutative group''.

@@

Mathematicians often ask, in which conditions the repeated application of an operation results in the neutral
element (compare patterns and structures before)\index{Structure}.

@@

\hypertarget{RSABeweis} Using the Euler-Fermat theorem, we can ``prove'' the
RSA\footnote{The RSA procedure\index{RSA procedure} is the most common asymmetric 
cryptography procedure. Developed in 1978 by Ronald Rivest, Adi Shamir and Leonard Adleman, it can be used both for
signatures and for encryption. Cryptographers always associate this procedure with the abbreviation ``RSA''. However, 
please note that the abbreviation may be used in other fields to mean
something completely different.}$^,$\footnote{%
In literature and movies not only classical methods are used (like the
secret message solved by Sherlock Holmes in ``The Adventure of the Dancing
Men'' by Arthur Conan  Doyle \index{Arthur Conan Doyle}), but also modern schemes. 
Examples are
% hier die items nicht einrücken ! xxxxxxxxxxxxxxx
\begin{list}{\textbullet}{\leftmargin10pt}
\item the movie ``House of Cards'', 1992, where autistic children
  communicate via primes,
\item the story ``The Dialogue of the Sisters'' by Dr.~C.~Elsner\index{Elsner 1999}, 1999
  (included in the CrypTool package\index{CrypTool} as PDF file), where the sisters
  communicate confidentially using a variant of RSA.
\end{list}\vspace{-\baselineskip}\vspace{-\parskip}}
procedure in the group $\mathbb{Z}_n^*$.

@@

Although Whitfield Diffie and Martin Hellman formulated the general method as
early as 1976, the actual procedure that met all four requirements was only
discovered later by Rivest, Shamir and Adleman.

@@

Since this is not true in general, we need to consider the case when $M$ and $p$ are not relatively prime. Since $p$ is a 
prime number, this implies that $p$ is a factor of $M$. But this means:
$$  M \equiv 0 {\rm ~(mod~}p). $$
If $p$ is a factor of $M$, then $p$ is also a factor of $M^{e * d}$. Therefore:
$$M^{e * d} \equiv 0 {\rm ~(mod~}p).$$
Since $p$ is a factor of both $M$ and $Me * d$, it is also a factor of their difference:
$$ (M^{e * d} - M ) \equiv 0 {\rm ~(mod~}p).$$
And therefore our conjecture is also true in this special case.

@@

{\bf 1st comment:} \\
We can also condense the three steps if we use the theorem \ref{thm-zth-fermateuler} (Euler-Fermat) - i.e. not the simplified theorem where $n = p$ and 
which corresponds to Fermat's Little Theorem:
$$
(M^e)^d \equiv M^{e*d} \equiv M^{(p-1)(q-1)*k + 1} \equiv 
        (\underbrace{M^{(p-1)(q-1)}}_{\equiv M^{J(n)} \equiv 1 {\rm ~(mod~}n)})^k * M
    \equiv 1^k * M \equiv M {\rm ~(mod~}n).
$$

@@

This made evident that a module length of 512 bit no longer prevents from
attackers.

@@

% --------------------------------------------------------------------------
\vskip +40pt
\subsubsection{Further current research in factorisation and number theory}
Prime numbers are part of very many topical research areas in number theory and computer science. 
Progress made with factorisation is bigger than estimated 5 years ago -- this
is not only founded on faster computers but also in new knowledge.

@@

Thus, any progress in efficiency of factorizing large integers will effect
the security of the RSA. As a consequence, the underlying primes $p$, $q$
and, thus, the module n (1024 bit as of today) have to be increased. In
the worst case of a breakthrough in factorization, the RSA algorithm might
be compromised.

% --------------------------------------------------------------------------
\vskip +20pt
\paragraph{Bernstein's paper and its implication on the
  security of the RSA algorithm}
\label{RSABernstein} \index{Factorisation problem}\mbox{}

@@

We note that the definition of \emph{effort} is a crucial point: Bernstein
claims that effort is the product of time and costs of the machine
(including the memory used). The gist of the paper lies in the fact that he
can reduce a big part of factorizing to sorting. Using Schimmler's scheme,
sorting can be optimized by massive parallel computing.  At the end of
section 3 Bernstein explains this effect: The costs of $m^2$ parallel
computers with a constant amount of memory is a constant times $m^2$.  The
costs of a computer with a single processor and memory of size $m^2$ is
also of the order of $m^2$, but with a different constant factor.  With
$m^2$ processors in parallel, sorting of $m^2$ numbers (with Schimmler's
scheme) can be achieved in time $m$, while a $m^2$-memory computer needs
time of the order of $m^2$. Decreasing memory and increasing the number of
processors, the computing time can be reduced by a factor $1/m$ without
additional effort in terms of total costs.  In section 5 it is said that
massive parallel computing can also increase efficiency of factorizing
using Lenstra's elliptic-curve-method (a search algorithm has costs that
increase quadratically instead of cubically).

We note that all results achieved so far are asymptotic results. This means
that they only hold in the limit n to infinity. Unfortunately, there is no
upper limit for the residual error (i.e. the difference between the real
and the asymptotic value) for finite n --- a problem which has already been
addressed by the author. As a consequence, one cannot conclude whether the
costs (in the sense of Bernstein) for factorizing 1024$-$2048-bit RSA modules
can be significantly reduced.

@@

When this section was written (June 2002) nothing was publicly known,
how far there exist implementations of his theoretical onsets and how
much financing there was for his research project.

@@

Their abstract summarizes their results very well: ``The security of the RSA
cryptosystem depends on the difficulty of factoring large integers. The best
current factoring algorithm is the Number Field Sieve (NFS), and its most
difficult part is the sieving step. In 1999 a large distributed computation
involving thousands of workstations working for many months managed to factor a
512-bit RSA key, but 1024-bit keys were believed to be safe for the next 15-20
years. In this paper we describe a new hardware implementation of the NFS
sieving step ... which is 3-4 orders of magnitude more cost effective than the
best previously published designs ... . Based on a detailed analysis of all the
critical components (but without an actual implementation), we believe that the
NFS sieving step for 1024-bit RSA keys can be completed in less than a year with
a \$10M device, and that the NFS sieving step for 512-bit RSA keys can be
completed in less than ten minutes with a \$10K device. Coupled with recent
results about the difficulty of the NFS matrix step ... this raises some
concerns about the security of these key sizes.''

@@

There are three tasks with completely different degrees of difficulty here. In
each case we know the cipher text 
and the public key $(e,n)$:
\begin{itemize}
\item[{\bf (a)}] Known plaintext: \index{Attack!known plaintext} find the secret key $d$ using the additionally known original message.
\item[{\bf (b)}] Cipher text only: \index{Attack!cipher text only} find $d$ and the plaintext.
\item[{\bf (c)}] Calculate the RSA modulus, in other words factorisation (with no knowledge of the message). \index{Factorisation problem}
\end{itemize}

@@

\bibitem[Eckert2003]{Eckert2003} \index{Eckert 2003}
    Claudia Eckert, \\
    {\em IT-Sicherheit: Konzepte-Verfahren-Protokolle}, 
    Oldenburg 2001, 2nd edition 2003.

@@

   \item \index{Munchenbach@M\"unchenbach}
         Only after I had completed this article, did I come across the
         website of Mr.\ M\"unchenbach, which interactively and didactically
         uses elementary number theory to provide a sophisticated
         description of the fundamental mathematical thought processes. 
         It was created for a teaching project in the 11th grade 
         of the technical grammar school (unfortunately only available in German): \\
         \href{http://www.hydrargyrum.de/kryptographie}
         {\texttt{http://www.hydrargyrum.de/kryptographie}}

    \item Once again only after finishing this I happened upon
          the web site of Mr.~Wagner, who is responsible for the development of
          the curriculum of computer science in one of the German federal 
          states (L\"ander). Here you can get hold of (but only in German) a collection of texts 
          and (Java-)\discretionary{}{}{}programs:\\
          \href{http://www.hom.saar.de/~awa/kryptolo.htm}
          {\texttt{http://www.hom.saar.de/\~{}awa/kryptolo.htm}}

@@

I would like to take this opportunity to thank 
\begin{itemize}
  \item {Mr.\ Henrik Koy for making many very useful suggestions, 
         for the very constructive proof-reading this article and
         for helping with TeX. \\
	 Mr.\ Koy developed the functions and the complex dialog box of the
	 RSA cryptosystem, which enable you to execute the RSA samples
	 of this article.}
  \item {J\"org-Cornelius Schneider for his TeX support und for the many 
         cases where he helped when facing programming or design problems.}
  \item {Dr. Illies for pointing me to Pari-GP \index{Pari-GP}.}
\end{itemize}

@@

Loops are also supported by Pari-GP. When you remove the table formatting
the result looks like this:
\begin{verbatim}
for( x=1,12, 
     for(y=1,12, print(Mod(x^y,m))); 
     if(gcd(x,m)==1, print(znorder(Mod(x,m))), print("--"));
     print(eulerphi(m)))
\end{verbatim}
